---
title: "Appendix 2: Data Analysis Documentation"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This appendix describes the R-code that was used for the analysis in the paper "Word-order variation in a contact setting: A corpus-based investigation of Russian spoken in Daghestan".


R version:
```{r}
getRversion()
```

Versions of the packages used in the analysis are specified at the very end of the document.


## Preparation of the data

Import the data in R
```{r, message=FALSE}
library("tidyverse")
```


```{r}
gen_std <- read.csv("std_rus.csv", stringsAsFactors=TRUE)
```


Set correct reference levels
```{r}
gen_std$head_lexical_class <- relevel(gen_std$head_lexical_class, "non_kinship")
gen_std$gen_lexical_class <- relevel(gen_std$gen_lexical_class, "non_human")
gen_std$gen_referentiality <- relevel(gen_std$gen_referentiality,"non_definite")
gen_std$gen_length <- relevel(gen_std$gen_length,"one-word")
```


## Random forest

```{r, message=FALSE}
library("party")
```

```{r, cache=TRUE}
gen_std_rf <- cforest(position ~ gen_lexical_class + head_lexical_class + 
                        gen_referentiality + gen_length,
                      data = gen_std, controls = cforest_unbiased(ntree = 1000, mtry = 2))
gen_std_varimp <- varimp(gen_std_rf, conditional = TRUE)
dotchart(sort(gen_std_varimp), main = "Conditional importance of variables")
```

```{r, message=FALSE}
library(Hmisc)
```


```{r}
gen_std_rf.pred <- unlist(treeresponse(gen_std_rf))[c(FALSE,TRUE)]
```

```{r}
somers2(gen_std_rf.pred, as.numeric(gen_std$position) -1)
```


```{r}
table(predict(gen_std_rf), gen_std$position)
```

```{r}
656/675
```



----

Versions of the packages used in the analysis:
```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
``` 



