---
title: "Appendix: Data Analysis Documentation"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

This appendix describes the R-code that was used for the analyses in the paper "Word-order variation in a contact setting: A corpus-based investigation of Russian spoken in Daghestan" by Chiara Naccarato, Anastasia Panova, Natalia Stoynova.

The R version:
```{r}
getRversion()
```

Versions of the packages used in the analysis are specified in the very end of the document.


## Data preparation

Set the working directory
```{r}
setwd("/Users/apanova/OneDrive/Documents/ConLab/WordOrder/WordOrder_R")
```


Import the data in R
```{r, message=FALSE}
library("tidyverse")
```

```{r}
gen <- read.csv("dag_rus.csv", stringsAsFactors=TRUE)
```

Set the correct reference levels
```{r}
gen$head_lexical_class <- relevel(gen$head_lexical_class, "non_kinship")
gen$gen_lexical_class <- relevel(gen$gen_lexical_class, "non_human")
gen$gender <- relevel(gen$gender,"m")
gen$gen_referentiality <- relevel(gen$gen_referentiality,"non_definite")
gen$gen_length <- relevel(gen$gen_length,"one-word")
gen$head_length <- relevel(gen$head_length,"multi-word")
gen$givenness <- relevel(gen$givenness,"other")
gen$year_of_birth <- relevel(gen$year_of_birth, "<1950")
```

```{r, include=FALSE}
#The symbol "more or equal" in the parameter `year_of_birth` creates an error during knitting to PDF, so it is replaced by the symbol "more".
levels(gen$year_of_birth) <- c("<1950", ">1950")
```


## Logistic regression


### Full model

```{r, message=FALSE}
library("lme4")
```

```{r}
model1 <- glmer (position ~ gen_lexical_class + head_lexical_class + education + 
                   gender + gen_referentiality + language_family + year_of_birth +
                   gen_individuation + gen_length + head_length + givenness + 
                   (1|speaker), data = gen, family ="binomial",
                 control = glmerControl(optimizer ="bobyqa"))
summary(model1)
```

### Step-wise selection procedure

```{r, warning=FALSE}
drop1(model1)
```

without year of birth
```{r, warning=FALSE}
model2 <- glmer (position ~ gen_lexical_class + head_lexical_class + education + 
                   gender + gen_referentiality + language_family + 
                   gen_individuation + gen_length + head_length + givenness + 
                   (1|speaker), data = gen, family ="binomial",
                 control = glmerControl(optimizer ="bobyqa"))
drop1(model2)
```

without givenness
```{r}
model3 <- glmer (position ~ gen_lexical_class + head_lexical_class + education + 
                   gender + gen_referentiality + language_family + 
                   gen_individuation + gen_length + head_length + 
                   (1|speaker), data = gen, family ="binomial",
                 control = glmerControl(optimizer ="bobyqa"))
drop1(model3)
```

without language family
```{r}
model4 <- glmer (position ~ gen_lexical_class + head_lexical_class + education + 
                   gender + gen_referentiality + 
                   gen_individuation + gen_length + head_length + 
                   (1|speaker), data = gen, family ="binomial",
                 control = glmerControl(optimizer ="bobyqa"))
drop1(model4)
```

without gen_individuation
```{r}
model5 <- glmer (position ~ gen_lexical_class + head_lexical_class + education + 
                   gender + gen_referentiality + 
                   gen_length + head_length + 
                   (1|speaker), data = gen, family ="binomial",
                 control = glmerControl(optimizer ="bobyqa"))
drop1(model5)
```

without education
```{r}
model6 <- glmer (position ~ gen_lexical_class + head_lexical_class + 
                   gender + gen_referentiality + 
                   gen_length + head_length + 
                   (1|speaker), data = gen, family ="binomial",
                 control = glmerControl(optimizer ="bobyqa"))
drop1(model6)
```

without head_length
```{r}
model7 <- glmer (position ~ gen_lexical_class + head_lexical_class + 
                   gender + gen_referentiality + 
                   gen_length + 
                   (1|speaker), data = gen, family ="binomial",
                 control = glmerControl(optimizer ="bobyqa"))
```

```{r}
drop1(model7)
```

### Minimal adequate model


```{r}
summary(model7)
```


Visualization of the estimates in the minimal adequate model

```{r, message=FALSE}
library(sjPlot)
library(ggplot2)
```

```{r}
plot_model(model7, type = "est", show.values = TRUE, sort.est = TRUE, value.offset = .3)
```


Obtaining the C value for the minimal adequate model

```{r, message=FALSE}
library(Hmisc)
```

```{r}
somers2(binomial()$linkinv(fitted(model7)), as.numeric(gen$position) -1)
```

Calculating the proportion of correctly predicted values

```{r, message=FALSE}
library("gmodels")
```

```{r}
fitted <- fitted(model7)
predicted <- ifelse(fitted >= .5, 1,0)
a <- data.frame(gen, predicted)
CrossTable(gen$position, a$predicted)
```

Investigating whether multicollinearity is a problem for the predictors in the model

```{r, message=FALSE}
library(languageR)
```

```{r, warning=FALSE}
collin.fnc(getME(model7, "X")[, -1])$cnumber
```


## Random forest

```{r, message=FALSE}
library("party")
```


parameters of the minimal adequate model

```{r}
gen_rf <- cforest(position ~ gen_lexical_class + head_lexical_class + 
                    gender + gen_referentiality + gen_length,
                  data = gen, controls = cforest_unbiased(ntree = 1000, mtry = 2))
gen_varimp <- varimp(gen_rf, conditional = TRUE)
dotchart(sort(gen_varimp), main = "Conditional importance of variables")
```



----

Versions of the packages used in the analysis:
```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
``` 

